---
title: 'Creating An Efficient Data Analysis Workflow : Book Sales Review'
author: "Eda AYDIN"
date: "28 03 2021"
output:
  pdf_document:
    toc: yes
---

# Upload necesessary libraries
```{r}
library(tidyverse)
library(lubridate)
```

# Data Preparation
```{r}
sales <- read_csv("sales2019.csv")
```
# Data Exploration
```{r}
# How big is the dataset? 
dim(sales)
```
```{r}
# What are the column names? What do they seem to represent?
colnames(sales)
```
The "date" column shows the data that the order of books was made.

```{r}
# What are the types of each of the columns?
for (col in colnames(sales)) {
  paste0(col, ":", typeof(sales[[col]])) %>% print
}
```
```{r}
# Do any of the columns have missing data?
for (col in colnames(sales)) {
  paste0(col, ", numbers of missing dara rows:",
          is.na(sales[[col]]) %>% sum) %>% print
}
```
The `user_submitted_review` column has some missing data in it. 

# Handling Missing Data
```{r}
# Remove the rows with no user_submitted_review
complete_sales <- sales %>%
  filter(
    !is.na(user_submitted_review)
  )
complete_sales
```
```{r}
# Calculate the mean of the total_purchased column, without the missing values

purchase_mean <- complete_sales %>%
  filter (!is.na(total_purchased)) %>%
  pull (total_purchased) %>%
  mean
purchase_mean
```
```{r}
# Assign this mean to all of the rows where total_purchased was NA
complete_sales <- complete_sales %>%
  mutate(
    imputed_purchased = if_else(is.na(total_purchased),
                                purchase_mean,
                                total_purchased)
  )
complete_sales
```
# Processing Review Data
```{r}
# Examine the unique sentences that are present in user_submitted_review

complete_sales %>% pull(user_submitted_review) %>% unique
```
```{r}
is_positive <- function(review) {
  review_positive = case_when(
    str_detect(review, "Awesome!") ~TRUE,
    str_detect(review, "Ok") ~ TRUE,
    str_detect(review, "a lot") ~TRUE,
    str_detect(review, "okay") ~ TRUE,
    str_detect(review, "Never") ~ TRUE,
    TRUE ~ FALSE # The review did not contain any of the above phrases
  )
}
```

```{r}
complete_sales <- complete_sales %>% 
  mutate(
    is_positive = unlist(map(user_submitted_review, is_positive))
  )
```

# Comparing Book Sales Between Pre- and Post- Program Sales
```{r}
complete_sales <- complete_sales %>%
  mutate (
    date_status = if_else(mdy(date) < ymd("2019/07/01"), "Pre", "Post")
  )
```

```{r}
complete_sales %>%
  group_by(date_status,title) %>%
  summarize(
    books_purchased = sum(imputed_purchased)
  ) %>%
  arrange(title, date_status)
```

# Comparing Book Sales Within Customer Type
```{r}
complete_sales %>%
  group_by(date_status, customer_type) %>%
  summarize (
    books_purchased = sum(imputed_purchased)
  ) %>%
  arrange(customer_type, date_status)
```

# Comparing Review Sentiment Between Pre- and Post-Program Sales
```{r}
# Create another summary table that compares the number of positive 
# reviews before and after July 1, 2019
complete_sales %>%
  group_by(date_status) %>%
  summarize(
    num_positive_reviews = sum(is_positive)
  )
```


