---
title: "Investigating COVID-19 Virus Trends"
author: "Eda AYDIN"
date: "21 02 2021"
output:
  pdf_document:
    toc: true
---

## Introduction

A pneumonia of unknown cause detected in Wuhan, China was first internationally reported from China on 31 December 2019. Today we know this virus as Coronavirus. COVID-19 which stands for COronaVIrus Disease is the disease caused by this virus. Since then, the world has been engaged in the fight against this pandemic. Several measures have therefore been taken to "flatten the curve". We have consequently experienced social distancing and many people have passed away as well.
```{r}

```

In the solidarity to face this unprecedented global crisis, several organizations did not hesitate to share several datasets allowing the conduction of several kinds of analysis in order to understand this pandemic.

It is natural for us to analyze these datasets by ourselves to answer questions since we cannot always rely on the news, and we are data scientists.

In this project, we use a dataset, from Kaggle. This dataset was collected between the 20th of January and the 1st of June 2020. The purpose of this Project is to build our skills and understanding of the data analysis workflow by evaluating the COVID-19 situation through this dataset. 

## Understanding Data 

```{r}
# importing the csv file
library(readr)
covid_df <- read_csv("covid19.csv")

```

```{r}
# finding dimension of data frame
dimension_ofdataset <- dim(covid_df)
print(dimension_ofdataset)
```

```{r}
# finding the number of columns in this data frame
ncol <- ncol(covid_df)
print(ncol)

# finding the number of rows in this data frame
nrow <- nrow(covid_df)
print(nrow)
```


```{r}
# finding the name of columns in this data frame
vector_cols <- colnames(covid_df)
print(vector_cols)
```

```{r}
# displaying the first six rows in this data frame
head_rows <- head(covid_df)
print(head_rows)

# displaying the last six rows in this data frame
tail_rows <- tail(covid_df)
print(tail_rows)
```

```{r}
# learning about a tibbles columns, types and dimensions
library(tibble)
glimpse(covid_df)
```

## Isolating the Rows We Need
```{r}
library(dplyr)
# filtering the rows related to "All States" from the Province_State
covid_df_all_states <- filter(covid_df, Province_State == "All States")
print(head(covid_df_all_states))

# removing the Province_State column from the data frame
covid_df <- select(covid_df, -Province_State)
print(head(covid_df))
```

## Isolating the Columns We Need
```{r}
# selecting the columns related to daily measures from covid_df_all_states
covid_df_all_states_daily <- select(covid_df_all_states, Date, Country_Region, active, 
                                    hospitalizedCurr, daily_tested, daily_positive)
print(head(covid_df_all_states_daily))
```

## Extracting the Top Ten Tested Cases Countries
```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarize(
  tested = sum(daily_tested),
  positive = sum(daily_positive),
  active = sum(active),
  hospitalized = sum(hospitalizedCurr)) %>% 
  arrange(-tested)

print(covid_df_all_states_daily_sum)
```

```{r}
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
print(covid_top_10)
```


## Identifying the Highest Positive Against Tested Cases
```{r}
# Creating the following vector from the covid_top_10 dataframe
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized
```

```{r}
# writing code to name the previous vectors by using names() function
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

```{r}
# identify the top three ratio
positive_tested_top_3 <- sort(positive_cases/tested_cases, decreasing = TRUE)
```

## Keeping relevant information

```{r}
# creating vectors
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)
```

```{r}
# creating matrix combining these vectors
covid_mat <- rbind(united_kingdom, united_states, turkey)

# rename the columns of this matrix with the vector
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
print(covid_mat)
```


## Pulling all together
```{r}
question <- "Which countries have had the highest number of 
positive cases against the number of tests?"

answer <- c("Positive tested cases" = positive_tested_top_3)
print(positive_tested_top_3[1:3])
```

```{r}
# creating list that contains the data structure
dataframes <- c(covid_df, covid_df_all_states, covid_df_all_states_daily, 
                covid_df_all_states_daily_sum, covid_top_10)

matrices <- covid_mat

vectors <- c(active_cases, countries, hospitalized_cases, 
             positive_cases, positive_tested_top_3)

data_structure_list <- c(dataframes, matrices, vectors)

covid_analysis_list <- c(question, answer, data_structure_list)
```

